Enriching Object Detection with 2D-3D Registration\\and Continuous Viewpoint Estimation

### 2015-04-01 intro.tex
minor change
grammar mod

\begin{figure}[t]
  \centering
  % \includegraphics[width=0.9\linewidth]{schematics} 
  \includegraphics[width=0.99\linewidth]{front} % \\[-5pt]
  % (a)\\[0pt]
  % \setlength\tabcolsep{0pt}
  % \begin{tabular}{cc}
  %     \includegraphics[width=0.45\linewidth]{car/orig2} & 
  %     \includegraphics[width=0.45\linewidth]{car/overlay2}
  %  %   \begin{tabular}{cc}
  %  %       \includegraphics[width=0.16\linewidth]{cad_car/2001-2004 Ford Ranger Edge} & 
  %  %       \includegraphics[width=0.16\linewidth]{cad_car/2002 Dodge Ram FedEx} \\ 
  %  %       \includegraphics[width=0.16\linewidth]{cad_car/2007-Nissan-Versa-_-Tiida-SL} & 
  %  %       \includegraphics[width=0.16\linewidth]{cad_car/2008-Jeep-Cherokee} \\ 
  %  %       \includegraphics[width=0.16\linewidth]{cad_car/2009 Toyota Cargo} & 
  %  %       \includegraphics[width=0.16\linewidth]{cad_car/2010-Chevrolet-Aveo5-LT} \\ 
  %  %       \includegraphics[width=0.16\linewidth]{cad_car/2012-Citroen-DS4}  
  %  %   \end{tabular}
  % \end{tabular}\\
  % (b)
  \caption{Using a database of 3D CAD models, we generate NZ-WHO
    templates which can be used to either detect objects directly or
    enrich the output of an existing detector with high-quality,
    continuous pose and 3D CAD model exemplar.}
    % initialize pose of an object in the bounding box from other
    % detectors. Once initialization is given, we use NZ-WHO to propose
    % and validate plausible pose on-the-fly and
    % further tune the translation, 3D viewpoint and focal length continuously.}
  % \vspace{-1em}
  \label{fig:front}
\end{figure}

@@

Current approaches to object class detection have reached a remarkable
level of performance in 2D bounding box
localization~\cite{pascal12, Felzenszwalb10, Overfeat, Krizhevsky12, Girshick14}, due to their
ability to generalize across differences in object appearance,
lighting, and viewpoint. While this generalization
ability of these methods is beneficial for robustness, it also limits
the level of detail that these detectors can deliver as an output.

As a consequence, there has been increased interest in multi-view object
recognition, where viewpoint estimates are provided by
detectors in addition to 2D bounding boxes. Several attempts have been
made to estimate viewpoint along with object class detections,
implemented as extensions to existing detectors, such as the implicit
shape models~\cite{sun10eccv}, the constellation model~\cite{stark10bmvc}, or the
deformable part model~\cite{Felzenszwalb10, gu10eccv,Xiang12,Pepik12,Fidler12,Hejrati14}.

Recently, an even higher level of geometric detail was reached in the
form of aligning
3D CAD model instances to real world test images~\cite{Aubry14,
  Lim14,Kholgade14, Chen13, Kostas14}. Interestingly, the problem of matching 3D models to 2D
images has been explored since the early days of computer
vision~\cite{Lowe87}, but had largely been neglected in recent years
in favor of 2D detectors based on robust local features and
statistical learning techniques. Now, this problem is being revisited
mostly due to two reasons: {\em (i)} the availability of 3D CAD models
for many object classes and {\em (ii)} the availability of robust
image matching techniques.
%
% performance object detectors usually generalizes very well. It
% can handle various intraclass variability, viewpoint change and
% occlusion. However due to this high generalization, sometimes it is
% very difficult to get accurate pose, category or 2D-3D matching. Our
% method, specialized in giving such accurate information, can be
% applied on any generic object detectors to give high quality metadata.
% %
% Matching 3D CAD model to an image has been a classical problem that
% dates back to eighties . Lowe \etal , the vision
% community instead focused on the object detection tasks that estimates
% a rectangular bounding box around an object\cite{Felzenszwalb10,
%   Girshick14}. 
%
%In addition, there has been a growing emand for accurate 2D-3D
%matching \cite{Kholgade14, Chen13, Kostas14} for image manipulation,
%editing, and data augmentation in the computer graphics community.
%
% Recently, with the powerful computing capabilities and a large
% collection of CAD models, the 2D-3D matching (registration) problem
% gets more and more approachable but due to inherent difficulty and
% limited performance of existing methods, many applications still
% require supervision to achieve 2D-3D matching.
% In medical community, registering 3D scan data to 2D image enabled various minimally invasive surgeries \cite{Markelj12}.
%
% For category-level matching, \cite{Xiang12, Hejrati14, Pepik12} used
% 3D parts or aspect to represent underlying category-level 3D geometry
% of the object. With the advent of large public CAD datasets
% \cite{Trimble, Yobi}, the instance-level object detection and 2D-3D
% registration became possible. \cite{Aubry14, Lim14} used large
% collection of CAD models to faithfully align 3D CAD model to an image
% by rendering CAD models and creating 2D templates for various
% viewpoints.

For (i), recent approaches to 2D-3D matching~\cite{Aubry14, Lim14}
typically rely on a
collection of 3D exemplar models, which they render from a
large number of viewpoints. The resulting artificial images are then
used to train exemplar models that can be matched to a real-world
image at test time.
%
For (ii), it has been realized that template-based exemplar detectors
based on HOG~\cite{Dalal05} features can be trained analytically, by
replacing the standard SVM with an LDA classifier
(E-LDA)~\cite{Hariharan12}, resulting in a whitened feature
representation, termed WHO. As a result, it is feasible to train
hundreds of thousands of mid-level patch detectors for
recognition~\cite{Aubry14}.
%
Unfortunately, the performance of WHO relies crucially on an
additional calibration step that equalizes the detection scores of
independently trained exemplar models. Since this step involves costly
mining of hard negative examples~\cite{Dalal05,Felzenszwalb10} on a
validation set, it constitutes the major computational bottleneck of
WHO, limiting its scalability.

In this paper, we propose a novel method for 2D-3D alignment of
exemplar CAD models to real-world images that circumvents the need for
calibration, and greatly enhances the scalability of WHO. As a
result, we can render novel views and train corresponding exemplar
models {\em on the fly}, without the need for offline processing. As a
by-product, we can formulate the alignment problem as a parameter
search in a continuous pose space, consisting of yaw, pitch and roll, resulting in truly continuous pose estimates, which
we implement using MCMC sampling.
%  require a large number of rendering and 2D
% templates for each of the CAD model with specific viewpoint. Since
% there are so many templates that it becomes almost impossible to get
% all the true positive image from training data. Instead \cite{Aubry14,
%   Lim14} used whitening as a preprocessing stage \cite{Hariharan12} to
% make templates more discriminative. Then calibration is done on a
% small training dataset.
%
% Although whitening to create template is faster than some of
% instance-level matching algorithms such as training Exemplar-SVM
% \cite{Malisiewicz11} or complicated models such as LEVAN
% \cite{Divvala14}, whitening and calibration large number of templates
% takes very long time. Also, once the fixed discrete viewpoint
% templates have been made and calibrated, it is difficult to fine-tune
% the prediction made using the large collection of templates.

Our paper makes the following specific contributions.

First, we present a novel method for training exemplar models from
rendered 3D CAD data {\em on the fly}, enabled by a novel variant of
WHO, termed NZ-WHO, and making efficient use of the specific
characteristics of rendered images. To our knowledge, our method
constitutes the first attempt to simultaneously render and train
exemplar detectors on the fly.
% In our paper, we present Non-Zero Whiten Histograms of Orientations
% (NZ-WHO) to dramatically speed up the template generation and get rid
% of time consuming calibration and yet achieves performance on par with
% calibrated templates. First, we will present NZ-WHO is more
% discriminative than WHO and introduce Conjugate Gradient method to
% speed up the whitening and create a high resolution template in just
% 80 milliseconds. Also, using this fast NZ-WHO template generation, we
% present \textbf{on-the-fly} template generation to fine tune an
% initial pose estimation to continuously refine translation, viewpoints
% (yaw, roll, pitch) and focal length.

Second, we show to enrich the output of an existing object class
detector, such as the DPM~\cite{Felzenszwalb10} or the
R-CNN~\cite{Girshick14} with additional 3D information, applying our
method to candidate detections provided by the respective detector. As
a result, we can refine the original detections with additional
estimates of 3D continuous pose and a 3D CAD model exemplar.

And third, we give an in-depth experimental study that demonstrates
the effectiveness of our approach on a standard benchmark for
object detection and viewpoint estimation~\cite{Xiang14}.

>>>>

\begin{figure}[t]
  \centering
  % \includegraphics[width=0.9\linewidth]{schematics} 
  \includegraphics[width=0.99\linewidth]{front} % \\[-5pt]
  \caption{Using a database of 3D CAD models, we generate NZ-WHO
    templates which can be used to either detect objects directly or
    enrich the output of an existing detector with high-quality,
    continuous pose and 3D CAD model exemplar.}
    % initialize pose of an object in the bounding box from other
    % detectors. Once initialization is given, we use NZ-WHO to propose
    % and validate plausible pose on-the-fly and
    % further tune the translation, 3D viewpoint and focal length continuously.}
  % \vspace{-1em}
  \label{fig:front}
\end{figure}

@@

Current approaches to object class detection have reached a remarkable
level of performance in 2D bounding box
localization~\cite{pascal12, Felzenszwalb10, Overfeat, Krizhevsky12, Girshick14}, due to their
ability to generalize across differences in object appearance,
lighting, and viewpoint. While this generalization
ability is beneficial for robustness, it limits
the level of detail that these detectors can deliver as an output.

As a consequence, there has been increased interest in multi-view object
recognition, where viewpoint estimates are provided by
detectors in addition to 2D bounding boxes. Several attempts have been
made to extend existing detectors to estimate viewpoint along with object class detections, including the implicit
shape model~\cite{sun10eccv}, the constellation model~\cite{stark10bmvc}, and the
deformable part model~\cite{Felzenszwalb10, gu10eccv,Xiang12,Pepik12,Fidler12,Hejrati14}.

Recently, an even higher level of geometric detail was reached in the
form of aligning
3D CAD model instances to real world test images~\cite{Aubry14,
  Lim14,Kholgade14, Chen13, Kostas14}. Interestingly, the problem of matching 3D models to 2D
images has been explored since the early days of computer
vision~\cite{Lowe87}, but had largely been neglected in recent years
in favor of 2D detectors based on robust local features and
statistical learning techniques. Now, this problem is being revisited
for two main reasons: {\em (i)} the availability of 3D CAD models
for many object classes and {\em (ii)} the availability of robust
image matching techniques.
%
% performance object detectors usually generalizes very well. It
% can handle various intraclass variability, viewpoint change and
% occlusion. However due to this high generalization, sometimes it is
% very difficult to get accurate pose, category or 2D-3D matching. Our
% method, specialized in giving such accurate information, can be
% applied on any generic object detectors to give high quality metadata.
% %
% Matching 3D CAD model to an image has been a classical problem that
% dates back to eighties . Lowe \etal , the vision
% community instead focused on the object detection tasks that estimates
% a rectangular bounding box around an object\cite{Felzenszwalb10,
%   Girshick14}. 
%
%In addition, there has been a growing emand for accurate 2D-3D
%matching \cite{Kholgade14, Chen13, Kostas14} for image manipulation,
%editing, and data augmentation in the computer graphics community.
%
% Recently, with the powerful computing capabilities and a large
% collection of CAD models, the 2D-3D matching (registration) problem
% gets more and more approachable but due to inherent difficulty and
% limited performance of existing methods, many applications still
% require supervision to achieve 2D-3D matching.
% In medical community, registering 3D scan data to 2D image enabled various
% minimally invasive surgeries \cite{Markelj12}.
%
% For category-level matching, \cite{Xiang12, Hejrati14, Pepik12} used
% 3D parts or aspect to represent underlying category-level 3D geometry
% of the object. With the advent of large public CAD datasets
% \cite{Trimble, Yobi}, the instance-level object detection and 2D-3D
% registration became possible. \cite{Aubry14, Lim14} used large
% collection of CAD models to faithfully align 3D CAD model to an image
% by rendering CAD models and creating 2D templates for various
% viewpoints.

For (i), recent approaches to 2D-3D matching~\cite{Aubry14, Lim14} rely on a
collection of 3D exemplar models, which they render from a
large number of viewpoints. The resulting artificial images are then
used to train exemplar models that can be matched to a real-world
image at test time.
%
For (ii), it has been realized that template-based exemplar detectors
based on HOG~\cite{Dalal05} features can be trained analytically, by
replacing the standard SVM with an LDA classifier
(E-LDA)~\cite{Hariharan12}. The result is a whitened feature
representation, termed WHO (Whitened Histogram of Orientations). This development makes it feasible to train
hundreds of thousands of mid-level patch detectors for
recognition~\cite{Aubry14}.
%
Unfortunately, the performance of WHO relies crucially on an
additional calibration step that equalizes the detection scores of
independently trained exemplar models. Since this step involves costly
mining of hard negative examples~\cite{Dalal05,Felzenszwalb10} on a
validation set, it constitutes the major computational bottleneck of
WHO, limiting its scalability.

In this paper, we propose a novel method for 2D-3D alignment of
exemplar CAD models to real-world images that circumvents the need for
calibration, and greatly enhances the scalability of WHO. As a
result, we can render novel views and train corresponding exemplar
models {\em on the fly}, without the need for offline processing. We call these NZ-WHO templates \ref{fig:front}. As a
by-product, we can formulate the alignment problem as a parameter
search in a continuous pose space, consisting of yaw, pitch and roll, which
we implement using MCMC sampling.

%  require a large number of rendering and 2D
% templates for each of the CAD model with specific viewpoint. Since
% there are so many templates that it becomes almost impossible to get
% all the true positive image from training data. Instead \cite{Aubry14,
%   Lim14} used whitening as a preprocessing stage \cite{Hariharan12} to
% make templates more discriminative. Then calibration is done on a
% small training dataset.
%
% Although whitening to create template is faster than some of
% instance-level matching algorithms such as training Exemplar-SVM
% \cite{Malisiewicz11} or complicated models such as LEVAN
% \cite{Divvala14}, whitening and calibration large number of templates
% takes very long time. Also, once the fixed discrete viewpoint
% templates have been made and calibrated, it is difficult to fine-tune
% the prediction made using the large collection of templates.

Our paper makes the following contributions:

First, we present a novel method for training exemplar models from
rendered 3D CAD data {\em on the fly}, enabled by a novel variant of
WHO, termed NZ-WHO, and making efficient use of the specific
characteristics of rendered images. To our knowledge, our method
constitutes the first attempt to simultaneously render and train
exemplar detectors on the fly.
% In our paper, we present Non-Zero Whiten Histograms of Orientations
% (NZ-WHO) to dramatically speed up the template generation and get rid
% of time consuming calibration and yet achieves performance on par with
% calibrated templates. First, we will present NZ-WHO is more
% discriminative than WHO and introduce Conjugate Gradient method to
% speed up the whitening and create a high resolution template in just
% 80 milliseconds. Also, using this fast NZ-WHO template generation, we
% present \textbf{on-the-fly} template generation to fine tune an
% initial pose estimation to continuously refine translation, viewpoints
% (yaw, roll, pitch) and focal length.
%
Second, we demonstrate that our method can enrich the output of an existing object class
detector, such as the DPM~\cite{Felzenszwalb10} or the
R-CNN~\cite{Girshick14} with additional 3D information. By applying our
method to candidate detections provided by the respective detector, we can augment the original detections with an
estimate of 3D continuous pose and a 3D CAD model exemplar.
%
Finally, we give an in-depth experimental study that demonstrates
the effectiveness of our approach on a standard benchmark for
object detection and viewpoint estimation~\cite{Xiang14}.

### 2015-04-01 mcmc.tex
grammar mod

This algorithm changes the current state $p$ to a new state $q$ based
on the acceptance probability
\begin{align}
    A(p \rightarrow q) & =  min\left( 1,  \frac{P(q) g(q \rightarrow p)}{P(p) g(p \rightarrow q)}\right).
\end{align}
We define $3$ different types of moves that can alter the state
(Fig.~\ref{fig:moves}), {\em (i)} changing the focal length $f$, {\em
(ii)} one of the rotational pose parameters $v_i$, and {\em (iii)} CAD model
index $m$. For (i) and (ii), we use Gaussian proposal distributions, and a uniform
distribution over model changes for (iii). For scale and translation, we
implicitly compute all possible translation and scale by convolving $w(q)$ to
HOG pyramid.
\begin{align}
    g(p \rightarrow p + v_i') & \sim \mathcal{N}(\mu = p_{v_i},\sigma = \sigma_v) \quad i \in \{1,2,3\}\\
    g(p \rightarrow p + f') & \sim \mathcal{N}(\mu = p_{f},\sigma = \sigma_f)\\
    g(p \rightarrow p + m') & \sim (1-c) \delta(p_m) + c Unif(1,M),
\end{align}
where $g$ is the proposal distribution. % and $P$ is the distribution of which we
%defined using Gaussian distribution around previous state. $A$ is the
%acceptance probability where we accept new proposal with $A(x \rightarrow x')$.
%
% One can also think of the image as an image cropped by detection
% bounding box. We synthesize covariance matrix and create NZ-WHO
% template $w(x)$ for particular viewpoint, focal length and CAD model.
% In sum, for new proposal $x'$, we make a NZ-WHO template on the fly 
% We found out that it requires good initialization to converge to
% local minima.

In practice, we run this algorithm for 20 iterations, keeping
the sample with the highest probability as our approximate estimate of
the MAP solution. We set $\sigma_v = 5$ degree and $\sigma_f = 1$ and
$c = 0.1$ for all experiments.

@@

\paragraph{Initialization.}
Since the objective function is non-convex, we need a good initialization in order to
increase the probability to find a solution that is close to the
global optima. We achieve this by first running a discrete,
pre-trained version of our algorithm (i.e., an ensemble of NZ-WHO
templates) in order to get promising candidate 2D bounding boxes and
poses to start from. We then initialize $x_0$ for each of these candidates.

>>>>

This algorithm changes the current state $p$ to a new state $q$ based
on the acceptance probability
\begin{align}
    A(p \rightarrow q) & =  min\left( 1,  \frac{P(q) g(q \rightarrow p)}{P(p) g(p \rightarrow q)}\right).
\end{align}
We define $3$ different types of moves that can alter the state
(Fig.~\ref{fig:moves}), {\em (i)} changing the focal length $f$, {\em
(ii)} one of the rotational pose parameters $v_i$, and {\em (iii)} CAD model
index $m$. For (i) and (ii), we use Gaussian proposal distributions, and a uniform
distribution over model changes for (iii). We implicitly compute all possible translation and scale by convolving $w(q)$ with a HOG pyramid.
\begin{align}
    g(p \rightarrow p + v_i') & \sim \mathcal{N}(\mu = p_{v_i},\sigma = \sigma_v) \quad i \in \{1,2,3\}\\
    g(p \rightarrow p + f') & \sim \mathcal{N}(\mu = p_{f},\sigma = \sigma_f)\\
    g(p \rightarrow p + m') & \sim (1-c) \delta(p_m) + c\,\textrm{Unif}(1,M),
\end{align}
where $g$ is the proposal distribution. % and $P$ is the distribution of which we
%defined using Gaussian distribution around previous state. $A$ is the
%acceptance probability where we accept new proposal with $A(x \rightarrow x')$.
%
% One can also think of the image as an image cropped by detection
% bounding box. We synthesize covariance matrix and create NZ-WHO
% template $w(x)$ for particular viewpoint, focal length and CAD model.
% In sum, for new proposal $x'$, we make a NZ-WHO template on the fly 
% We found out that it requires good initialization to converge to
% local minima.

In practice, we run this algorithm for 20 iterations, keeping
the sample with the highest probability as our approximate estimate of
the MAP solution. We set $\sigma_v = 5^\circ$, $\sigma_f = 1$ and
$c = 0.1$ for all experiments.

@@

\paragraph{Initialization.}
Since the objective function is non-convex, we need a good initialization in order to
increase the probability that we find a solution close to the
global optima. We achieve this by first running a discrete,
pre-trained version of our algorithm (i.e., an ensemble of NZ-WHO
templates) in order to get promising candidate 2D bounding boxes and
poses to start from. We then initialize $x_0$ for each of these candidates.
